---
title: "Gun Violence in America"
author: "Sri Gautham Subramani"
output: 
  revealjs::revealjs_presentation:
    theme: league
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(viridis)
library(shiny)
library(revealjs)
library(leaflet)
library(ggmap)
library(htmltools)
library(date)
library(lubridate)
library(htmlwidgets)
library(tidyverse)
library(choroplethr)
library(choroplethrMaps)
library(repmis)
library(ggthemes)
library(sp)
library(base)
library(jpeg)
library(openintro)
library(rgdal)
library(raster)
library(mapview)
```

##Objective

>- The Objective of this topic is to discuss the severity of Gun Violence in America
>- To address the Statistics behind major shootings
>- To render the results visually with the help of R

## Data Sources

>- [Gun violence database](https://www.kaggle.com/gunviolencearchive/gun-violence-database/data) : Discussing incidents of gun violence from 2014-2016.
>- [World firearms murders and ownership](https://docs.google.com/spreadsheets/d/1chqUZHuY6cXYrRYkuE0uwXisGaYvr7durZHJhpLGycs/edit#gid=0) : Discussing the total number of population of gun owners in each country
>- [Stanford Data](https://raw.githubusercontent.com/StanfordGeospatialCenter/MSA/master/Data/Stanford_MSA_Database.csv?raw=True) : Discussing some of the important shootings from 1966-2013

##Setting up the Data

>- Gun Violence Data was not readily available.The main website [Gun Violence Archive](http://www.gunviolencearchive.org/) although is open to everyone to view the data, downloads are restricted only to the first 500 entries.
>- The other difficulties was to combine the yearwise data to one complete dataset.
>- To format the date and extract the Year to perform Statistical Analysis.
>- Some unique datasets were taken from multiple websites through **Web Scraping**

##Initial Analysis



```{r loadData, echo=FALSE, warning=FALSE, message=FALSE}
 mass_shootings_2014 <- source_data("https://github.com/DataWrangling957/HomeworkData/blob/master/Final%20Project/mass_shootings_2014.csv?raw=True")
 mass_shootings_2015 <- source_data("https://github.com/DataWrangling957/HomeworkData/blob/master/Final%20Project/mass_shootings_2015.csv?raw=True")
 mass_shootings_2016 <- source_data("https://github.com/DataWrangling957/HomeworkData/blob/master/Final%20Project/mass_shootings_2016.csv?raw=True")
 stanford_data <- source_data("https://raw.githubusercontent.com/StanfordGeospatialCenter/MSA/master/Data/Stanford_MSA_Database.csv?raw=True")
 full_data <- source_data("https://github.com/DataWrangling957/HomeworkData/blob/master/Final%20Project/full_data.csv?raw=True")
 world_data <- source_data("https://raw.githubusercontent.com/DataWrangling957/HomeworkData/master/Final%20Project/World%20firearms%20murders%20and%20ownership%20-%20Sheet%201.csv?raw=True")
 mass_shootings <- rbind(mass_shootings_2014, mass_shootings_2015, mass_shootings_2016)
```

```{r echo=TRUE, include=TRUE, warning=FALSE, message=FALSE}
mass_shootings$Incident.Date <- mdy(mass_shootings$`Incident Date`)
mass_shootings <- mass_shootings %>% mutate(Year = substring(.$Incident.Date,1,4))
mass_shootings$State <- casefold(mass_shootings$State, upper = FALSE)
state_demographics <- mass_shootings %>% group_by(State) %>% summarise(n = sum(`# Killed`)) %>% ungroup()
state_demographics %>% arrange(-n)
```

## State Choroplethr

```{r, warning=FALSE}
state_demographics %>% 
  dplyr::rename(region = State, value = n) %>%
  state_choropleth(title = "Number of Mass Shootings", num_colors = 1)
```

## Examining Stanford Data

```{r, warning=FALSE}
stanford_data$Date <- mdy(stanford_data$Date)
stanford_data <- stanford_data %>% mutate(Year = substring(.$Date,1,4))

stanford_demographics <- 
  stanford_data %>% group_by(State) %>% summarise(n = sum(`Total Number of Victims`)) %>% ungroup() %>% arrange(-n)
stanford_demographics$State <- casefold(stanford_demographics$State, upper = FALSE)
stanford_demographics %>% 
  dplyr::rename(region = State, value = n) %>%
  state_choropleth(title = "Number of Shootings", num_colors = 1)
```

## Gun Violence by Year of Stanford Data

```{r, echo=TRUE, warning=FALSE}
stanford_data %>% group_by(Year) %>% summarise(n = sum(`Total Number of Victims`)) %>% ungroup() %>% ggplot(mapping = aes(Year,n,color = Year,fill = Year)) + 
  geom_bar(stat = "identity") + theme_calc() + scale_fill_discrete(na.value = "some colour other than grey")
```

## Gun Violence by Place Type

```{r, echo=TRUE, warning=FALSE}
stanford_data %>% group_by(`Place Type`) %>% summarise(n = sum(`Total Number of Victims`)) %>% ungroup() %>% arrange(-n)
```

## Gun Violence by Age

```{r, echo=TRUE, warning=FALSE}
stanford_data %>% group_by(`Shooter Age(s)`) %>% summarise(n = sum(`Total Number of Victims`)) %>% ungroup() %>% arrange(-n)
```

## Gun Violence Based on Intent

```{r, echo=TRUE, warning=FALSE}
full_data_updated <- mutate(full_data, Count = rep(1,nrow(full_data)))
deaths_data <- full_data_updated %>% group_by(intent) %>% summarise(n = sum(Count)) %>%  ungroup() %>% arrange(-n)
deaths_data %>% na.omit()
```

## Gun Deaths based on Ethinicity

```{r, echo=TRUE, warning=FALSE}
stanford_data %>% group_by(`Shooter Race`) %>% summarise(n = sum(`Total Number of Victims`)) %>% ungroup() %>% arrange(-n)
```

##Analysis of Global Gun Owners
```{r echo=FALSE, warning=FALSE, message=FALSE}
df_country <- world_data[,c(1,9)]
names(df_country) <- c("region","value")
df_country$region <- casefold(df_country$region , upper = FALSE)
df_country$region[176] <- c("united states of america")
df_country$region[39] <- c("republic of congo")
df_country$region[53] <- c("united kingdom")
df_country$region[44] <- c("northern cyprus")
df_country$region[166] <- c("united republic of tanzania")
df_country$region[89] <- c("north korea")
df_country$region[90] <- c("south korea")
df_country$region[147] <- c("republic of serbia")
df_country$region[39] <- c("democratic republic of the congo")
```

```{r, echo=TRUE, warning=FALSE}
country_choropleth(df_country)
world_data <- as.data.frame(world_data)
```

##Population of Gun Owners
```{r, echo=FALSE, warning=FALSE}
highest_firearms <- world_data %>% group_by(`Country/Territory`) %>% summarise(Total = sum(`Average total all civilian firearms`)) %>% ungroup() %>% arrange(-Total) %>% head(n=10)

ggplot(data = highest_firearms, mapping = aes(highest_firearms$`Country/Territory`, Total, colour = highest_firearms$`Country/Territory`, size = highest_firearms$Total)) + geom_point()
```